---
title: "Tournament Importance Plots"
author: "Inki Sul"
date: "December 17, 2015"
output: pdf_document
---

Golf Project (Embarked: 9/23/2015)
Tournament Importance Plots for Tiger Woods
Last Updated: 12/17/2015
Author:  Professor Chaithanya Bandi, Inki Sul

Focusing on Tiger Woods, our favorite exemplary player, as always.

```{r}
require(e1071)
require(stringr)
library(reshape2)
require(ggplot2)
require(gridExtra)
library(grid)
library(gtable)
require(UsingR)
if(!require(coefplot)) { install.packages("coefplot"); require(coefplot)}
if(!require(doBy)) { install.packages("doBy"); require(doBy)}
if(!require(plyr)) { install.packages("plyr"); require(plyr)}
if(!require(dplyr)) { install.packages("dplyr"); require(dplyr)}
require(rpart)
require(rpart.plot)
library(randomForest)
library(tree)
library(party)
if(!require(ROCR)) { install.packages("ROCR"); require(ROCR)}
if(!require(forecast)) { install.packages("forecast"); require(forecast)}
if(!require(zoo)) { install.packages("zoo"); require(zoo)}
if(!require(neuralnet)) { install.packages("neuralnet"); require(neuralnet)}

### Functions to be used later:
# Reordering function within dataframe for prettier data
moveMe <- function(data, tomove, where = "last", ba = NULL) {
  temp <- setdiff(names(data), tomove)
  x <- switch(
    where,
    first = data[c(tomove, temp)],
    last = data[c(temp, tomove)],
    before = {
      if (is.null(ba)) stop("must specify ba column")
      if (length(ba) > 1) stop("ba must be a single character string")
      data[append(temp, values = tomove, after = (match(ba, temp)-1))]
    },
    after = {
      if (is.null(ba)) stop("must specify ba column")
      if (length(ba) > 1) stop("ba must be a single character string")
      data[append(temp, values = tomove, after = (match(ba, temp)))]
    })
  return(x)
}
# Function used for plotting a histogram of forecast errors, with an overlaid N(0, SD) curve:
plotForecastErrors <- function(forecasterrors)
{
  # make a histogram of the forecast errors:
  mybinsize <- IQR(forecasterrors)/4
  mysd   <- sd(forecasterrors)
  mymin  <- min(forecasterrors) - mysd*5
  mymax  <- max(forecasterrors) + mysd*3
  # generate normally distributed data with mean 0 and standard deviation mysd
  mynorm <- rnorm(10000, mean=0, sd=mysd)
  mymin2 <- min(mynorm)
  mymax2 <- max(mynorm)
  if (mymin2 < mymin) { mymin <- mymin2 }
  if (mymax2 > mymax) { mymax <- mymax2 }
  # make a red histogram of the forecast errors, with the normally distributed data overlaid:
  mybins <- seq(mymin, mymax, mybinsize)
  hist(forecasterrors, col="red", freq=FALSE, breaks=mybins)
  # freq=FALSE ensures the area under the histogram = 1
  # generate normally distributed data with mean 0 and standard deviation mysd
  myhist <- hist(mynorm, plot=FALSE, breaks=mybins)
  # plot the normal curve as a blue line on top of the histogram of forecast errors:
  points(myhist$mids, myhist$density, type="l", col="blue", lwd=2)
}
getwd()
setwd("C:/Users/Inki Sul/Documents/Current Documents/Research/Golf/raw-data")

# Read in event level raw data + tournament importance
data <- read.csv('event-level-raw-comparisonTest-importance.csv',header=TRUE, sep=",",row.names=NULL)
# Subsetting the data for Tiger Woods and its performance variables
tiger <- {
  d <- subset(data[grep("Tiger", data$Player.First.Name, ignore.case = TRUE, fixed = FALSE),], select=c(1,2,3,4,8,9,10,14,195,197,118,78,79))
  d <- d[order(d$Tournament.Year,d$Tournament.Number),]
  d$Tournament.Order <- 1:nrow(unique(data.frame(x=d$Tournament.Year, y=d$Tournament.Number)))
  d
}
tiger$SG_Total <- tiger$SG_T2G + tiger$SG_Putts
```

Now, take a look at the original Strokes Gained plot for Tiger Woods.
```{r, echo=FALSE}
SG <- ggplot(melt(subset(tiger, select=c(14,9,10)), id=1), aes(Tournament.Order,value,fill=variable)) + 
  geom_bar(position="dodge",stat="identity") + 
  theme(legend.justification = c(0, 1), legend.position = c(0, 1))
print(SG)
```

What follow are the Strokes Gained and its compenents plotted against tournament order numbered chronologically, or against player age.
```{r, echo=FALSE}
SGT1 <- ggplot(tiger, aes(x=Tournament.Order,y=SG_Total,group=Tournament.Importance)) + 
  geom_line(aes(colour=Tournament.Importance)) + 
  theme(legend.justification = c(0, 1), legend.position = c(0, 1))
print(SGT1)

SGT2 <- ggplot(tiger, aes(x=Player.Age.Numeric,y=SG_Total,group=Tournament.Importance)) + 
  geom_line(aes(colour=Tournament.Importance)) + 
  theme(legend.justification = c(0, 1), legend.position = c(0, 1))
print(SGT2)

SGPutt1 <- ggplot(tiger, aes(x=Tournament.Order,y=SG_Putts,group=Tournament.Importance)) + 
  geom_line(aes(colour=Tournament.Importance)) + 
  theme(legend.justification = c(0, 1), legend.position = c(0, 1))
print(SGPutt1)

SGPutt2 <- ggplot(tiger, aes(x=Player.Age.Numeric,y=SG_Putts,group=Tournament.Importance)) + 
  geom_line(aes(colour=Tournament.Importance)) + 
  theme(legend.justification = c(0, 1), legend.position = c(0, 1))
print(SGPutt2)

SGT2G1 <- ggplot(tiger, aes(x=Tournament.Order,y=SG_T2G,group=Tournament.Importance)) + 
  geom_line(aes(colour=Tournament.Importance)) + 
  theme(legend.justification = c(0, 1), legend.position = c(0, 1))
print(SGT2G1)

SGT2G2 <- ggplot(tiger, aes(x=Player.Age.Numeric,y=SG_T2G,group=Tournament.Importance)) + 
  geom_line(aes(colour=Tournament.Importance)) + 
  theme(legend.justification = c(0, 1), legend.position = c(0, 1))
print(SGT2G2)
```

We can also take a look at some other key performance metric: Greens in Regulation Percentage as well. (No zero dropped as NA)

```{r, echo=FALSE}
GIR_original <- ggplot(tiger, aes(x=Tournament.Order,y=Total_GIR)) + geom_line()
print(GIR_original)

GIR_importance1 <- ggplot(tiger, aes(x=Tournament.Order,y=Total_GIR,group=Tournament.Importance)) + 
  geom_line(aes(colour=Tournament.Importance)) + 
  theme(legend.justification = c(0, 1), legend.position = c(0, 1))
print(GIR_importance1)

GIR_importance2 <- ggplot(tiger, aes(x=Player.Age.Numeric,y=Total_GIR,group=Tournament.Importance)) + 
  geom_line(aes(colour=Tournament.Importance)) + 
  theme(legend.justification = c(0, 1), legend.position = c(0, 1))
print(GIR_importance2)
```

More to follow.....